---
title: "Nursery analysis"
author: "DPIRD - RP"
format: html
---

```{r setup}
#| echo: false

# load libs
library(arrow)
library(data.table)
library(dplyr)
library(ggplot2)
library(lubridate)
library(mgcv)
library(readxl)

# source functions
source("R/utils.R")

# Get all file names
all_files <- list.files("data/NR"); all_files

# Grab only ones with average temperature
temperature_files <- grep("_avgTemp10mins_", all_files, value = T)

# Behaviour index
behaviour_index_file <- grep("_hourlyActivityIndex_", all_files, value = TRUE)

# Mortality files
sick_files <- grep("_Sickpen_Mortality.", all_files, value = TRUE)

# Pen and sex metadata
tag_id_sex_file <- grep("_taggingIds_", all_files, value = TRUE)
```


## Process files and export as parquet for fast(er) I/O

```{r}
# Read in all temperature files
tag_temp_df <- rbindlist(
  lapply(temperature_files, 
         function(x) fread(paste0("data/NR/", x))
  )
)

# Run a few checks
str(tag_temp_df)
summary(tag_temp_df)

  # Unique tag
n_pigs <- length(unique(tag_temp_df$ANIMAL_ID)); n_pigs

# set names
names(tag_temp_df) <- c("tag_id", "tag_temp", "date_time")

# Add day_night column based on hour column
tag_temp_df[, hour := as.integer(format(date_time, "%H"))]
tag_temp_df[, day_night := fifelse(hour >= 18 | hour < 6, 0, fifelse(hour >= 6 | hour < 18, 1, NA_integer_))]

# Set theoretical temperature thresholds
tag_temp_df[, hypo_crit := fifelse(tag_temp <= 35 & tag_temp > 32, 1, 0)]
tag_temp_df[, hypo_severe := fifelse(tag_temp <= 32, 1, 0)]

# Fever
tag_temp_df[, fever_crit := fifelse(tag_temp >= 40 & tag_temp < 42, 1, 0)]
tag_temp_df[, fever_severe := fifelse(tag_temp >= 42, 1, 0)]

# Optimum
tag_temp_df[, optimum_temp := fifelse(tag_temp >= 38.5 & tag_temp <= 39.5, 1, 0)]

# Export data as parquet
arrow::write_parquet(tag_temp_df, "data/processed/batch2_tag_temp.parquet")

# Behaviour index data ----
# Read in
index_df <- fread(paste0("data/NR/", behaviour_index_file))
str(index_df)
summary(index_df)

names(index_df) <- c("tag_id", "activity", "date_time")

# Add day_night column based on hour column
index_df[, hour := as.integer(format(date_time, "%H"))]
index_df[, day_night := fifelse(hour >= 18 | hour < 6, 0, fifelse(hour >= 6 | hour < 18, 1, NA_integer_))]

# Export files as parquet for easier I/O
arrow::write_parquet(index_df, "data/processed/batch2_behaviour_index.parquet")

# Mortality data ----
sick_list <- 
  lapply(sick_files, 
         function(x) fread(paste0("data/NR/", x))
  )

sick_df <- 
  sick_list[[2]] |>
  dplyr::select(Tag_ID, Sex, `Mortality date`, `Final Activity Value`) |>
  janitor::clean_names()

sick_df <- sick_df[1:36, ]  

str(sick_df)
summary(sick_df)

# Manually fix the date for tag #0184
sick_df[35, ]
sick_df[35, "mortality_date"] <- "2-Oct"

# Place NA in empty tag_id rows
sick_df[3, ]$tag_id <- NA
sick_df[18, ]$tag_id <- NA

# parse dates properly formatted
dates <- paste0(sick_df$mortality_date, "-2024")
dates <- lubridate::parse_date_time(dates, "dmy", tz = "Australia/Perth")
sick_df$mortality_date <- dates

sick_df <-
  dplyr::filter(sick_df, !is.na(tag_id))

# Export to processed dir
arrow::write_parquet(sick_df, "data/processed/batch2_sick_record.parquet")

# Pen data ----
# Get a list of all the worksheet names
sheet_names <- readxl::excel_sheets(file.path("data/NR", tag_id_sex_file))

# Loop through each worksheet
pen_data_list <- list()

counter <- 1
for (sheet in sheet_names) {
  # Read the worksheet into a data.table
  
  dt <- 
    as.data.table(
      read_xlsx(file.path("data/NR", tag_id_sex_file), 
                sheet = sheet)
      ) 
  
  dt <- janitor::clean_names(dt)

  dt$tag_id <- 
    substring(dt[["xiot_tag_id"]], nchar(dt[["xiot_tag_id"]]) - 3, nchar(dt[["xiot_tag_id"]])) 
  
  pen_data_list[[counter]] <- dt
  counter <- counter + 1
}

pen_data_list[[22]]$pen_number <- 22

names(pen_data_list[[23]])[1] <- "pen_number"
names(pen_data_list[[24]])[1] <- "pen_number"

# Select the pen#, tag_id and gender columns
pen_data_df_clean <- 
  lapply(
    pen_data_list,
    extract_data_from_pen_data
  ) |>
  rbindlist() |>
  dplyr::mutate(tag_id = factor(tag_id),
                gender = factor(fifelse(gender == "M", 0, 1, NA)),
                pen_number = factor(pen_number))

# Export to processed dir
arrow::write_parquet(pen_data_df_clean, "data/processed/batch2_pen_gender_ids.parquet")

```

## Nursery data

We are importing the pre-processed data for the pigs in the **nursery pen 2**. These files are the tag temperature measurements aggregated every 10 min over a 24h period. The period is then split into day (6am to 6pm) and night (6pm to 6 am). (NOTE check times). There are `r length(list.files("data/NR"))` files, one for each week of the **batch 2** collection (eg *`r list.files("data/NR")[1]`*). 

```{r check_data_eda}

# Load datasets
tag_temp_df <- arrow::read_parquet("data/processed/batch2_tag_temp.parquet")
behaviour_idx_df <- arrow::read_parquet("data/processed/batch2_behaviour_index.parquet")
sick_df <- arrow::read_parquet("data/processed/batch2_sick_record.parquet")
pen_gender_df <- arrow::read_parquet("data/processed/batch2_pen_gender_ids.parquet")

# Match sick with pen data
sick_complete_df <- 
  sick_df |>
  left_join(pen_gender_df, by = "tag_id")

sick_complete_df$sex <- NULL
names(sick_complete_df)[4] <- "sex"

# Check observations for each tag (10 mins over time period - if complete x obs)
# Temperature data
obs_counts_temp <- 
  tag_temp_df |> 
  dplyr::mutate(tag_id = factor(tag_id)) |>
  group_by(tag_id) |>
  count() |>
  arrange(-n)

hist(obs_counts_temp$n)
max(obs_counts_temp$n) # 3994

# Match temp counts with pen metadata
tag_temp_metadata_df <- 
  obs_counts_temp |>
  left_join(pen_gender_df, by = "tag_id")

# 
obs_counts_idx <- 
  behaviour_idx_df |> 
  dplyr::mutate(tag_id = factor(tag_id)) |>
  group_by(tag_id) |>
  count() |>
  arrange(-n)

max(obs_counts_idx$n) # 613
hist(obs_counts_idx$n)

# Match temp counts with pen metadata
tag_idx_metadata_df <- 
  obs_counts_idx |>
  left_join(pen_gender_df, by = "tag_id")

# Check for outliers or obvious errors
# Get tag_id with missing observations
obs_counts_temp[which(obs_counts_temp$n < 2000), ] |> arrange(n)

# Only tags FF78, 9CB5, 029C, FEDC seem to have issues (perhaps tags of dead pigs?)
#  1 FF78       4
#  2 9CB5      23
#  3 029C      71
#  4 FEDC     120
#  5 02C4     140
#  6 004B     140
#  7 0017     141
#  8 0111     141
#  9 01B0     141
# 10 0268     141

tags_with_missing_obs <- 
  filter(obs_counts_temp, n < 2000) |>
  dplyr::pull(tag_id) |>
  as.character()
  
# FF78 (only 4 obs, removed)
filter(new_df, tag_id == "FF78")

# 9CB5 (only 23 obs, removed), they seem extreme, and measurements missing:
# between 1:30 and 2:40 am then from 4:10 to 8:10 am (last obs)
filter(new_df, tag_id == "9CB5")

# 029C (71 obs, removed) they seem extreme, and measurements missing mostly 
# during the morning/day and a gaps overnight, has a 200°C observation
filter(new_df, tag_id == "029C")

# FEDC (FEDC obs, no clear issues, only missing a obs at the end)
filter(new_df, tag_id == "FEDC")

# check with a plot
# add some of the tags with full obs
df_missing_obs <- 
  filter(tag_temp_df, tag_id %in% tags_with_missing_obs) |> #c("FF78", "9CB5", "029C", "FEDC", "0000", "0003", "0002", "0004", "0005", "0007")) |>
  group_by(tag_id) |>
  arrange(date_time) |>
  ungroup() |>
  setDT() |>
  dplyr::mutate(tag_id = factor(tag_id))

p1_missing_tag_obs <-
  ggplot(df_missing_obs, aes(x = date_time, y = tag_temp), group = tag_id) +
  geom_line(aes(colour = tag_id), linewidth = 1.6, alpha = .8) +
  # scale_color_manual(values = c(rep("grey", 6), "firebrick", "orange", rep("firebrick", 2))) +
  scale_x_datetime() +
  scale_y_continuous(limits = c(15, 60), breaks = seq(25, 50, 5)) +
  theme_minimal(base_size = 16)


# Filter out tags with issues
eda_df <- 
  new_df |>
  filter(tag_id %notin% c("FF78", "9CB5", "029C"))

eda_df[, condition := fifelse(fever_crit == 1 | hypo_crit, "alert", 
                              fifelse(fever_severe == 1 | hypo_severe == 1, "sick",
                                      fifelse(optimum_temp == 1, "optimal", "normal")))]

eda_df <-
  eda_df |> mutate(condition = factor(condition),
                   day_night = factor(day_night))

tags <- unique(levels(eda_df$tag_id))

ggplot(subset(eda_df, tag_id == c("FF9E", "01CD", "DF4A")), 
       aes(x = dates, y = tag_temp, group = tag_id, colour = tag_id)) +
  geom_line(linewidth = 1.2, alpha = .8) +
  scale_x_datetime() +
  scale_y_continuous(limits = c(15, 55), breaks = seq(15, 55, 5)) +
  # facet_wrap(. ~ day_night) +
  theme_minimal(base_size = 16)




x <- get_temperature_stats(eda_df, group = "all")
stats_by_tag <- get_temperature_stats(eda_df, group = "tag_id")

# Set theoretical temperature thresholds
stats_by_tag <- setDT(stats_by_tag)

stats_by_tag[, hypo_crit := fifelse(avg <= 35 & avg > 32, 1, 0)]
stats_by_tag[, hypo_severe := fifelse(avg <= 32, 1, 0)]

# Fever
stats_by_tag[, fever_crit := fifelse(avg >= 40 & avg < 42, 1, 0)]
stats_by_tag[, fever_severe := fifelse(avg >= 42, 1, 0)]

# Optimum
stats_by_tag[, optimum_temp := fifelse(avg >= 38.5 & avg <= 39.5, 1, 0)]

stats_by_tag[, condition := fifelse(fever_crit == 1 | hypo_crit, "alert", 
                                    fifelse(fever_severe == 1 | hypo_severe == 1, "sick",
                                            fifelse(optimum_temp == 1, "optimal", "normal")))]


t <- subset(stats_by_tag, sd < 2)

stats_by_tag[, threshold := (avg + (0.5 * sd))]

test_df <-
  eda_df |>
  filter(tag_id %in% my_df$tag_id)
  # filter(tag_id %in% c(sample(tags, 10, replace = F)))

ggplot(test_df, aes(x = tag_temp)) + 
  geom_density() +
  facet_wrap(. ~ day_night) +
  theme_minimal()

ggplot(t, aes(x = tag_id, y = avg, group = day_night, fill = day_night)) +
  geom_col(alpha = .8) +
  scale_color_viridis_d() +
  # scale_x_datetime() +
  # scale_y_continuous(limits = c(15, 60), breaks = seq(25, 50, 5)) +
  facet_wrap(. ~ day_night) +
  theme_minimal(base_size = 16)





eda_df <- 
  new_df |> 
  group_by(tag_id) |>
  mutate(obs_count = n()) |>
  ungroup()

filter(eda_df, obs_count < 144) |>
  group_by(tag_id) |>
  distinct()

ggplot(new_df, aes(x = dates, y = tag_temp), group = day_night) + 
  geom_point() + 
  scale_x_datetime() + 
  theme_minimal() 

filter(new_df, tag_temp > 47)

# If `tag_temp` over critical threshold: alert
# If `tag_temp` over severe threshold (<32 or >40): sick

new_df[, condition := fifelse(fever_crit == 1 | hypo_crit, "alert", 
                              fifelse(fever_severe == 1 | hypo_severe == 1, "sick",
                                      fifelse(optimum_temp == 1, "optimal", "normal")))]

str(new_df)

# new_df |>
#   dplyr::mutate(condition2 = factor(condition2)) |>
#   dplyr::group_by(day_night, condition2) |>
#   dplyr::reframe(avg_temp = mean(tag_temp, na.rm = T))
# 
# 
# new_df[, condition2 := ifelse((hypo_severe + fever_severe) == 2, "sick", 
#                               fifelse((hypo_severe + fever_severe) == 1, "alert", "normal"))]
# 
# 
# x <- group_split(new_df |> group_by(day_night))
# 
# x[[1]] |>
#   group_by(tag_id) 
# 
# 
# # How many records for each pig this week
# record_counts <- new_df[, .(RecordCount = .N), by = tag_id]; record_counts
# meanT <- new_df[, .(temp_avg = mean(tag_temp)), by = tag_id]
# mean_sd <- new_df[, .(temp_sd = sd(tag_temp)), by = tag_id]
# 
# 
# x <- merge(record_counts, meanT)
# x <- merge(x, mean_sd)
# 
# custom_temp_up <- function(avg, sd) {return(avg + (0.5 * sd))}
# custom_temp_down <- function(avg, sd) {return(avg - (0.5 * sd))}
# 
# 
# # Get data for one pig
# pig_FFFD <- subset(new_df, tag_id == "0000")
# pig_FFFD$rolling_avg <- zoo::rollmean(pig_FFFD$tag_temp, 6, fill = NA)
# pig_FFFD$rolling_sd <- zoo::rollapply(pig_FFFD$tag_temp, 6, fill = NA, "sd")
# 
# 
# 
# pig_FFFD |>
#   ggplot(aes(x = dates, y = tag_temp)) +
#   geom_line(alpha = .6) +
#   geom_line(aes(x = dates, y = rolling_avg), colour = "firebrick") +
#   # scale_y_continuous(limits = c(28, 42)) +
#   scale_x_datetime(date_breaks = "3 hours") +
#   theme_minimal()
# 
# 
# fit <- lm(tag_temp ~ as.character(tag_int), data = new_df)
# fit
# summary(fit)
# plot(fit)
# 
# lapply(coefs, add_num, intercept)





```
